---
title: "HW2_STAT210P_MinhCao"
output:
  pdf_document: default
  html_document: default
date: "2025-01-31"
editor_options:
  markdown:
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Say a regression equation if fit to data, and the correlation coefficient estimate ,R ,between X and Y is 0.5.

State if true or false.

a.The slope of the regression line is 0.5: False\

b.The regression model with X explains 50% of the variation in Y: False 

c.25% of the variation in Y is explained by X: True\

d.50% of the variation in Y is explained by X: False (R\^2 = 0.25: coefficient of determination) (R = 0.5: Correlation Coefficient)\

e.  X is positively associated with Y: True\

f.Even if X and Y have a non linear relationship, the value of 0.5 can be used to measure the association between X and Y.: False

## 2 Suppose a simple linear model is fit to predict Y = weight in kilograms using X= height in centimeters of an adult. But say a new simple linear model is fit using Y = height in centimeters and X= weight in kilograms (that is to say Y and X have reversed).State whether each of the following would be the same for this new model as it was for the original model, or it would be different and explain in a sentence or two.

a.  The value of R.: Stay the same as the formula stay the same  

b.  The value of R2: Stay the same as value of R stay the same  

c.  The estimate of B1: Different, In the original model, B1 represents the change in Weight per unit change in Height.
    In the new model, it represents the change in Height per unit change in Weight.\

d.  The estimate of B0: Different.
    The intercept B0 represents the expected value of the dependent variable when the independent variable is zero\

e.  The test statistic to test if the correlation between explanatory and response is equal to 0.: Same because correlation stays the same

## 3

```{r}
mid_path = "C:/Users/caoqu/OneDrive/Desktop/UCI/STUDY/WINTER_25/DATA_210P_Statistical_Methods_I/data/MidwestSales.txt"
Midwest = read.table(mid_path, sep= "", header = FALSE)
head(Midwest)

```

```{r}
names(Midwest)=c("id","price","sqft","bed","bath","ac","garage","pool","year","quality",
"style","lot","hwy")
```

a.  Fit a linear model where the square footage of the house is used to predict the sale price. (X=sqft and Y =price). Write the estimated regression equation

```{r}
model1 = lm(price ~ sqft, data = Midwest)
summary(model1)
```

-   Estimated regression equation: Price = -81432.946 + 158.950\*sqft

b.  Interpret the estimate of the slope.

When average sqft (X) increases by 1 unit, we expect the price of the house to increase by \$158.950

c. Test whether square footage has a significant linear relationship with price. Use alpha = 0.05 level of significance. Write the null and alternative hypothesis, state the test statistic and p-value, and make a conclusion

R^2 = 0.6715 and B1 = 158.950 \

```{r}
# Therefore R = 

R = 0.6715**0.5
R
```


- H0: p = 0 \
- Ha: p != 0 \ 

- Test statistics t* = 
```{r}
# Test statistics t*
df = 520
t_star = (R * sqrt(df - 2)) / (sqrt(1 - R**2))
t_star
```
```{r}
# Finding p-value

p_value = 2*pt(q = t_star, df = 518, lower.tail = FALSE)
p_value
```
Using alpha = 0.05 and p_value = 2.589982e-127, we reject null hypothesis and accept alternative hypothesis. We conclude that square footage has a significant linear relationship with price


d. Now test whether square footage has a significant positive linear relationship with price. Use alpha = 0.05 level of significance. Write the null and alternative hypothesis, state the test statistic and p-value, and make a conclusion.

Same as c, we reject null hypothesis and accept alternate hypothesis, and since B1 > 0, we conclude that square footage has a significant positive linear relationship with price

e.Find and interpret a 95% confidence interval for the mean price when sqft=2000 \
```{r}
# 95% confidence interval

predict(model1, list(sqft = 2000), interval = "c" )
```

f.Find and interpret a 95% prediction interval for the price when sqft=2000

```{r}
# 95% prediction interval
predict(model1, list(sqft = 2000), interval = "p" )

```

g. What would happen to the interval from part f. if the confidence level is decreased to 90%? Explain in a sentence or two

The interval will be smaller as the multiplier become smaller

h.Would it make sense to predict the sale price of a house that is 8500 square feet?

Yes


## 4.This question will use the skin cancer data set that is on the class website. Say it is of interest if latitude is predictive of mortality rate due to skin cancer. Fit a simple linear model where X=Lat and Y =Mort.

```{r}
skin_path = "C:/Users/caoqu/OneDrive/Desktop/UCI/STUDY/WINTER_25/DATA_210P_Statistical_Methods_I/data/skincancer.txt"
skin = read.table(skin_path, sep= "", header = TRUE)
head(skin)
```

```{r}
model2 = lm(Mort ~ Lat, data = skin)
summary(model2)
```



a. Using a alpha = 0.05 significance level, conduct a formal statistical test of whether latitude has a linear association with the mortality.

R^2 = 0.6715 and B1 = 158.950 \

```{r}
# Therefore R = 

R = 0.6798**0.5
R
```


- H0: p = 0 \
- Ha: p != 0 \ 

- Test statistics t* = 
```{r}
# Test statistics t*
df = 47
t_star = (R * sqrt(df - 2)) / (sqrt(1 - R**2))
t_star
```
```{r}
# Finding p-value

p_value = 2*pt(q = t_star, df = 45, lower.tail = FALSE)
p_value
```
Using alpha = 0.05 and p_value = 1.057931e-12, we reject null hypothesis and accept alternative hypothesis. We conclude that latitude has a linear association with the mortality.

b. Find and interpret a 99% confidence interval for the mean mortality rate when Lat=40.

```{r}
# 99% confidence interval
predict(model2, list(Lat = 40), interval = "c" )
```

c. Find and interpret a 99% prediction interval for the mortality rate when Lat=40.
```{r}
# 99% prediction interval
predict(model2, list(Lat = 40), interval = "p" )
```

d. What can you say about the center of the confidence interval and prediction interval. Is it the same?

Yes they are the same. Both intervals are centered at the same point—the predicted value from your model

e. How does the width of the confidence interval compare to the prediction interval. Explain in a sentence or two.

The prediction interval is wilder than the confidence interval because it includes the extra uncertainty from the variability of individual observations, not just the uncertainty in estimating the mean


## 5 Remember the form of the prediction interval. The standard error of the prediction at a value of X = Xp was SE( Y^p)

For each part, state what will happen to the prediction interval (stay the same, be wider, or be
narrower).

a. The sample size is increased.: This make the prediction interval narrower
b. If Xp gets closer to X bar, the average of the X covariate: This make the prediction interval narrower
c. Wider
d. Stay the same

